---
title: "Problem Set 3"
author: "Lindsey Cathcart and Rose Pfeiffer"
date: "10/2/2020"
output: html_document
bibliography: ProblemSet3Bib.bib
csl: apa.csl

---

# Species of concern: Bog Fritillary Butterfly (*Procolossiana eunomia*)

##### 1. 
We chose to study this species due to concerns around its conservation.
Like most other insects, this is a poorly studied species. There are
indications that loss of bog habitat vital to the growth of its host
plants has negative consequences on the butterfly population
[@Goffart2010]. Climate change is also thought to play a role as increased
temperatures have various effects on life stages [@Radchuk2013]. The
species has been observed to be absent from several areas from which it
historically would have been found in, including Armenia [@Morgun2019].
Specifically, we are interested in the population dynamics in Belgium, as
the population there is isolated from other Northern European populations
due to lack of proper habitat between the populations [@Goffart2010].


##### 2. 
 We would like to answer the question, "Which life stage of the Bog Fritillary butterfly should be considered most important for conservation?"


##### 3.
Calculate eigenvalue, stable age distribution, elasticity, and sensitivity. What does this tell you about the population?

```{r, echo=FALSE}

#filter out the specific species
# comadre$metadata 
# grep searches within a vector for a word, so search genus for proclossiama
# grep(pattern = 'Proclossiana',x=comadre$metadata$Genus)
# head(comadre)
# my_matrix<-comadre$mat[1176]
# head(my_matrix)

#install.packages("demogR")
library("demogR")

# build a matrix
A <- matrix(c(0,0,0,0,89,0.7444,0,0,0,0,0,0.49,0,0,0,0,0,0.025,0,0,0,0,0,0.75,0), nr = 5, byrow=TRUE)
```


```{r}
A
# calculate eigenvalue
my_eigen <- eigen(A)
my_eigen$values[5]

# calculate stable age dist.
eigen.analysis(A)
stable_age <- eigen.analysis(A)$stable.age
stable_age

# calculate sensitivities
sensitivities <- eigen.analysis(A)$sensitivities
sensitivities

# calculate elasticities
elasticities <- eigen.analysis(A)$elasticities
elasticities

```
The eigen value (`r my_eigen$values[5]`) indicates that the population is declining as it is less than 1. 
The stable age distribution (`r stable_age`), is the right eigen vector.
It tells us that for the population to be stable, the majority of
individuals should be in the 1st or 2nd life stages, egg and pre-diapausal
larvae. Approximately 20% of individuals should be in the overwintering
larvae stage. Only about 1% of the population should be in the pupa and
adult stages for the population to be stable. 
The sensitivity matrix, shows us that the third life stage, overwintering
larvae, has the greatest sensitivity (`r sensitivities[4,3]`). This
indicates that a small change to the survival rate at this stage would
have the greatest impact on the population. The elasticities were equal
for all life stages.

##### 4. 
From the sensitivity calculation it is clear that implementing
conservation strategies at the overwintering larval stage would provide
the greatest increase in population growth. 


The stable age distribution shows us that the greatest losses are occurring in the 'overwintering larva: active' stage, where the observed survival probability was 0.025, but the stable age distribution shows that a survival probability of 0.19432 is needed to maintain a stable distribution.


how to interpret the stable age distribution values...?

Increased temperature --> reduces overwintering survival, which is why the sensitivity is so high then...


## Notes
sensitivity vs elasticity
sensitivity: how elements in vital rate (vital rate refers to survivorship at stage 1, fecundity at stage 2, etc) effect the population growth rate. which age/stage is most receptive to control?

elasticity is a proportional change in vital rate

both tell you how the change in rates (growth/fecundity/survival) affect the overall growth rate.  

the stable age distribution is the proportion of the population in each age or stage class, and mathematically is represented by the right eigenvector. the distribution of the stages is stable, not the growth/decline of the population. the distribution

easton's github: eastonwhite

##### Works Cited


















