---
title: "Problem Set 3"
author: "Lindsey Cathcart and Rose Pfeiffer"
date: "10/2/2020"
output: html_document
bibliography: ProblemSet3Bib.bib
csl: apa.csl

---

# Species of concern: Bog Fritillary (*Procolossiana eunomia*)


1. We chose to study this species due to concerns around its conservation. Like most other insects, this is a poorly studied species. There are indications that loss of bog habitat vital to the growth of its host plants has negative consequences on the butterfly population [@Goffart2010]. Climate change is also thought to play a role as increased temperatures have various effects on life stages [@Radchuk2013]. The species has been observed to be absent from several areas from which it historically would have been found in, including Armenia [@Morgun2019]. Specifically, we are interested in the population dynamics in Belgium, as the population there is isolated from other Northern European populations due to lack of proper habitat between the populations [@Goffart2010].


2. Which life stage should be considered most important in conservation?


3. Calculate eigenvalue, stable age distribution, elasticity, and sensitivity. What does this tell you about the population?

```{r, echo=FALSE}

#filter out the specific species
# comadre$metadata 
# grep searches within a vector for a word, so search genus for proclossiama
# grep(pattern = 'Proclossiana',x=comadre$metadata$Genus)
# head(comadre)
# my_matrix<-comadre$mat[1176]
# head(my_matrix)

#install.packages("demogR")
library("demogR")

# build a matrix
A <- matrix(c(0,0,0,0,89,0.7444,0,0,0,0,0,0.49,0,0,0,0,0,0.025,0,0,0,0,0,0.75,0), nr = 5, byrow=TRUE)
```


```{r}
A
# calculate eigenvalue
my_eigen <- eigen(A)
my_eigen$values[5]

# calculate stable age dist.
eigen.analysis(A)
stable_age <- eigen.analysis(A)$stable.age
stable_age

# calculate sensitivities
sensitivities <- eigen.analysis(A)$sensitivities
sensitivities

# calculate elasticities
elasticities <- eigen.analysis(A)$elasticities
elasticities

```


The stable age distribution shows us that the greatest losses are occurring in the 'overwintering larva: active' stage, where the observed survival probability was 0.025, but the stable age distribution shows that a survival probability of 0.19432 is needed to maintain a stable distribution.

how to interpret the stable age distribution values...?

Increased temperature --> reduces overwintering survival, which is why the sensitivity is so high then...

sensitivity vs elasticity
sensitivity: how elements in vital rate (vital rate refers to survivorship at stage 1, fecundity at stage 2, etc) effect the population growth rate. which age/stage is most receptive to control?

elasticity is a proportional change in vital rate

both tell you how the change in rates (growth/fecundity/survival) affect the overall growth rate.  

the stable age distribution is the proportion of the population in each age or stage class, and mathematically is represented by the right eigenvector. the distribution of the stages is stable, not the growth/decline of the population. the distribution

easton's github: eastonwhite



















