---
title: "Problem Set 3"
author: "Lindsey Cathcart and Rose Pfeiffer"
date: "10/2/2020"
output: html_document
---

# Species of concern: *Procolossiana eunomia*


1. We chose to study this species due to concerns around its conservation. Like most other insects, this is a poorly studied species, although there are indications that loss of bog habitat vital to the growth of its host plants has negative consequences on the butterfly population (citation?). It has been observed to be absent from several areas from which it originally would have been found in, including Armenia (citation?). Specifically we are interested in the population dynamics in Belgium.


2. Which life stage should be considered most important in conservation?


3. Calculate eigenvalue, stable age distribution, elasticity, and sensitivity. What does this tell you about the population?

```{r, echo=FALSE}

#filter out the specific species
# comadre$metadata 
# grep searches within a vector for a word, so search genus for proclossiama
# grep(pattern = 'Proclossiana',x=comadre$metadata$Genus)
# head(comadre)
# my_matrix<-comadre$mat[1176]
# head(my_matrix)

#install.packages("demogR")
library("demogR")

# build a matrix
A <- matrix(c(0,0,0,0,89,0.7444,0,0,0,0,0,0.49,0,0,0,0,0,0.025,0,0,0,0,0,0.75,0), nr = 5, byrow=TRUE)
```


```{r}
A
# calculate eigenvalue
my_eigen <- eigen(A)
my_eigen$values[5]

# calculate stable age dist.
eigen.analysis(A)
stable_age <- eigen.analysis(A)$stable.age
stable_age

# calculate sensitivities
sensitivities <- eigen.analysis(A)$sensitivities
sensitivities

# calculate elasticities
elasticities <- eigen.analysis(A)$elasticities
elasticities

```


The stable age distribution shows us that the greatest losses are occurring in the 'overwintering larva: active' stage, where the observed survival probability was 0.025, but the stable age distribution shows that a survival probability of 0.19432 is needed to maintain a stable distribution.

how to interpret the stable age distribution values...?

Increased temperature --> reduces overwintering survival, which is why the sensitivity is so high then...