---
title: "Problem Set 3"
author: "Lindsey Cathcart and Rose Pfeiffer"
date: "10/2/2020"
output: html_document
bibliography: ProblemSet3Bib.bib
csl: apa.csl

---

# Species of concern: Bog Fritillary Butterfly (*Procolossiana eunomia*)

##### 1. 
We chose to study this species due to concerns around its conservation.
Like most other insects, the population dynamics of this species are poorly understood. However, there are
indications that loss of bog habitat vital to the growth of its host
plants has negative consequences on the butterfly population
[@Goffart2010]. Climate change is also thought to play a role on declining populations, as increased
temperatures have various effects on life stages [@Radchuk2013]. The
species has been observed to be absent from several areas from which it
historically would have been found in [@Morgun2019].
Specifically, we are interested in the population dynamics in Belgium, as
the population there is isolated from other Northern European populations
due to lack of habitat that would allow for dispersal between them [@Goffart2010].


##### 2. 
 We would like to answer the question, "Which life stage of the Bog Fritillary butterfly should be considered most important for conservation?"


##### 3.
Calculate eigenvalue, stable age distribution, elasticity, and sensitivity. What do these tell you about the population?

```{r, echo=FALSE}

#filter out the specific species
# comadre$metadata 
# grep searches within a vector for a word, so search genus for proclossiama
# grep(pattern = 'Proclossiana',x=comadre$metadata$Genus)
# head(comadre)
# my_matrix<-comadre$mat[1176]
# head(my_matrix)

#install.packages("demogR")
library("demogR")

# build a matrix
A <- matrix(c(0,0,0,0,89,0.7444,0,0,0,0,0,0.49,0,0,0,0,0,0.025,0,0,0,0,0,0.75,0), nr = 5, byrow=TRUE)
```


```{r}
A
# calculate eigenvalue
my_eigen <- eigen(A)
my_eigen$values[5]

# calculate stable age dist.
eigen.analysis(A)
stable_age <- eigen.analysis(A)$stable.age
stable_age

# calculate sensitivities
sensitivities <- eigen.analysis(A)$sensitivities
sensitivities

# calculate elasticities
elasticities <- eigen.analysis(A)$elasticities
elasticities

```
The eigen value (`r my_eigen$values[5]`) indicates that the population is declining as it is less than 1. 
The stable age distribution (`r stable_age`), is the right eigen vector.
It tells us that for the distribution of ages in the population to be stable, the majority of
individuals should be in the 1st or 2nd life stages (egg and pre-diapausal
larvae). Approximately 20% of individuals should be in the overwintering
larvae stage. Only about 1% of the population should be in the pupa and
adult stages for the population to be stable. At those frequencies, the distribution of ages in the population is stable.  
The sensitivity matrix shows us that the third life stage, overwintering
larvae, has the greatest sensitivity (`r sensitivities[4,3]`). This
indicates that a small change to the survival rate at this stage would
have the greatest impact on the population. The elasticities for each vital rate were equal to 0.2, representing the proportional change in overall growth rate from changes in vital rates (such as fecundity or survivorship).



##### 4. 
Based on the sensitivites calculated, it is clear that implementing
conservation strategies at the overwintering larval stage would provide
the greatest increase in population growth. That stage is very sensitive, so to increase the eigenvalue to be greater than 1, an increase in the survival at this life stage is necessary. These strategies may include rearing additional larvae under controlled climate and habitat restoration to assure a proper microclimate for overwintering.






##### Works Cited


















